---
title: "p8105_hw3_rd3096"
output: github_document
---

# load necessary pakages 
```{r setup}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)
data("ny_noaa")
```

## problem 1

### read and explore the dataset "ny_noaa"

we explore the dataset and get a summary of the dataset

```{r}
str(ny_noaa)
summary(ny_noaa)
```

### we write a short description of the dataset and answer the following questions

Our dataset ahs `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns.It includes key variables such as id (station identifier), date, prcp (precipitation), snow (snowfall), snwd (snow depth), tmax (maximum temperature), and tmin (minimum temperature)."
The variable types based on our output are listed as follows:

* id: character

* date: Date format

* prcp, snow, snwd: integer

* tmax, tmin: character (we wil convert these into numeric later for proper analysis).

Missing data is a significant issue in the ny_noaa dataset, especially for snow depth (snwd) and snowfall (snow), Precipitation has less missing data, but it still affects a non-negligible portion of the dataset.

we will do some data cleaning for subquestion 1
```{r}

ny_noaa_cleaned <- ny_noaa %>%
  mutate(
    year = as.numeric(format(date, "%Y")),
    month = as.numeric(format(date, "%m")),
    day = as.numeric(format(date, "%d"))
  )

```



